Brownian motion on an oscillating system

### 2017-06-13 examples/oscillator.Rmd
.

# Brownian motion

Now suppose that $A$ is changing due to mutation, 
and that it takes the value $A(t)$ after $t$ units of evolutionary time.
We know that $A(t)$ must equal $A_\tau$ for some $\tau$, 
and hence we can define $A(t) = A_{W(t)}$, 
where $W(t)$ is a random process on the free parameter;
but what are the appropriate dynamics for $W$?
Suppose that genetic drift acts uniformly on the entries of $A$, 
so that ...

# Toy example

As a toy example, suppose that we want to describe a process through $\R^2$
on the set
$$\begin{aligned}
    x^2 + y^2 = 1,
\end{aligned}$$
i.e., the unit circle.
This is parameterized by
$$\begin{aligned}
    x(\theta) &= \cos(\theta) \\
    y(\theta) &= \sin(\theta) .
\end{aligned}$$

The dynamics we want are limits of the following discrete dynamics.
Let $\epsilon$ denote the step size,
and suppose that at rate $\eta$ steps of size $\epsilon$ 
in a uniformly random direction $\phi$ are proposed;
but the probability that such a step is accepted is Kimura's expression:
$$\begin{aligned}
    q(x',y';x,y) = \frac{1 - e^{-2 (f(x',y')-f(x,y))}}{1 - e^{-2 N_e (f(x',y')-f(x,y))}}
\end{aligned}$$
where $(x',y')$ is the new (proposed) location, and
$$\begin{aligned}
    f(x,y) = (\sqrt{x^2+y^2}-1)^2
\end{aligned}$$
is the fitness function.
Note that $q(x,y;x,y) = 1$ 
and that 
$$\begin{aligned}
    \partial_u q(u,v;x,y) 
    &= 
        ( \partial_u f(u,v) ) 
        \left( \partial_z \frac{1 - e^{-2 (z-f(x,y))}}{1 - e^{-2 N_e (z-f(x,y))}} \vert_{z=f(x',y')} \right) \\
    &= 
        \left( \frac{2 x (\sqrt{x^2+y^2}-1)}{\sqrt{x^2+y^2}} \right) 
        \left( \frac{ 2 e^{-2 (f(x',y')-f(x,y))} (1 - e^{-2 N_e (f(x',y')-f(x,y))}) - 2 N_e e^{-2 (f(x',y')-f(x,y))} }{(1 - e^{-2 N_e (f(x',y')-f(x,y))})^2} \right)
\end{aligned}$$
and so that
$$\begin{aligned}
    \partial_u q(u,v;x,y) \vert_{u=x, v=y}
    &= 
        \left( \frac{2 x (\sqrt{x^2+y^2}-1)}{\sqrt{x^2+y^2}} \right) 
        \left( \frac{ - 2 N_e  }{(1 - e^{-2 N_e (f(x',y')-f(x,y))})^2} \right)
\end{aligned}$$

This has the result of multiplying jump rates by the acceptance probabilities;
so equivalently, if the process is at $(x,y)$,
then it jumps to 
$$\begin{aligned}
    (x',y') = (x+\epsilon \cos \phi, y + \epsilon \sin \phi)
\end{aligned}$$
at rate
$$\begin{aligned}
    \eta q(x',y';x,y) 
    &\approx

\end{aligned}$$

>>>>

We can define the dynamics in two ways: extrinsically or intrinsically.
Extrinsically,
we look at the dynamics of a population mean in $\R^{2\times 2}$,
thinking about $A$ as evolving.
Intrinsically, we move about only on the set of equivalent $A$,
letting $\tau$ evolve.
As the population size goes to infinity,
the extrinsic dynamics should approach the intrinsic dynamics.

# Extrinsic dynamics

Let $N$ denote the population size,
and suppose that at any time the population is spread about from the mean $A$
as a Gaussian with variance $\sigma^2$, independently in each entry of $A$.
Then we expect that the population mean takes steps of size $\sigma/\sqrt{N}$
with a restoring force towards the set of optimal $A$.
